---
title: "Demonstration of GFL-NMR via the `gflnma` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GFL-NMR Example: Atrial Fibrillation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gflnma)
library(netmeta)
```

# Demonstration with a single covariate

## Data

To illustrate GFL-NMR, we will make use of a network of evidence on 17 
stroke prevention treatments for those with atrial fibrillation [Cooper 2009, 
van Valkenhoef 2023]. There are 25 studies in this data set. The outcome is the 
number of patients who experienced a stroke, and we will model the rate of 
patients who experienced a stroke, which is calculated as event/exposure, where 
event = 'number of patients who experienced a stroke' and exposure = 'total 
patient years at risk'. The relative effect measure (`TE`) is the log rate ratio.

The covariate we will adjust for is proportion of patients with prior stroke 
(`stroke`).

Let's examine the data more closely. Note that, since the `gflnma` R package 
relies on the `netmeta` R package to compute the between-study standard 
deviation in random effects models, all $\binom{T_m}{2}$ pairwise contrasts in a 
multi-arm trial $m$ comparing $T_m$ treatments must be inputted (e.g., 3 
contrasts for 3-arm trials, 6 contrast for 4-arm trials, etc.).

```{r}
head(dat_afib, n = 12)
```

The are several multi-arm trials in this data set. As an example, AFASAK 1 is
a 3-arm trial and thus we have three contrasts. AFASAK 2 is a 4-arm trial and 
thus we have six contrasts.

The data set contains arm-level (`event1`, `time1`, `event2`, `time2`) and 
contrast-level data (`TE`, `seTE`). While GFL-NMR requires contrast-level data,
we have included arm-level data which will later be used to compute the 
covariance between the relative effects in multi-arm trials.

### Inspect network

We can call upon `netconnection()` from the `netmeta` R package to assess if the 
network is connected as well as `netgraph()`, also from the `netmeta` R package, 
to plot the network diagram.

```{r, fig.dim = c(6, 6)}
# Check to see if it is connected
(netcon_check <- netmeta::netconnection(dat_afib, 
                                        treat1 = treat1, 
                                        treat2 = treat2, 
                                        studlab = studlab))

# Plot network diagram
netmeta::netgraph(netcon_check)
```

The output from `netmeta::netconnection()` tells us there is only one 
network, meaning there is a single connected network.

## Derive inputs for GFL-NMR

To implement the GFL-NMR framework, we require 1) a data frame that provides 
contrast-level data on relative treatment effects vs. a trial-specific reference 
treatment estimated in each study, and 2) the corresponding variance-covariance 
matrix. `prep_gfl_data()` will derive these outputs based on an inputted data 
set that contains all pairwise contrasts in multi-arm studies. 

The outputted variance-covariance matrix depends on whether a fixed, random, or
multiplicative effects model is desired, which must be specified by the 
`modtype` argument. See note at the end of this vignette for more details.

```{r}
dat_list <- prep_gfl_data(dat = dat_afib, 
                          TE = "TE",
                          seTE = "seTE", 
                          treatment1 = "treat1", 
                          treatment2 = "treat2",
                          studlab = "studlab",
                          event1 = "event1", 
                          time1 = "time1",
                          modtype = "ME")

head(dat <- dat_list$dat)

head(var_cov <- dat_list$var_cov)

(mod_type <- dat_list$mod_type)
```

Notice how the data in the multi-arm studies now only contain evidence on the 
relative treatment effects vs. a study-specific reference treatment. 

We can then extract the study-level summary statistics for the covariates we 
would like to adjust for. Here, we will adjust for the proportion of patients 
who previously experienced a stroke, which is saved in the `stroke` column. Note 
that the covariate variables will be internally scaled by `solve_gflnma()` when 
we later solve the GFL-NMR problem(s).

```{r}
# extract the study-level covariate data
x_cov <- as.matrix(dat$stroke)
```

## Single problem

We will first demonstrate GFL-NMR based on a single value for `gamma`. This 
means there is only one GFL-NMR problem to solve. 

### Solve GFL-NMA problem: gamma = 1

Now that we have the required data inputs, we may solve the GFL-NMR problem via
the `solve_gflnma()` function. Within `solve_gflnma()`, there are some arguments
that are specific to NMR: `class1`, `class2`, and `ref_class`, which may be used 
to pre-specify some desired groupings for the relative treatment effect-covariate
interaction terms (e.g., because there is not enough evidence for one 
interaction term to be independently estimated. GFL-NMR, however, can group 
treatment-covariate interaction terms when this is the case, so the user need 
not do this); `gamma`, which is the ratio of the penalty term for the 
interaction terms ($\beta_{k,p}$) over the penalty term for the relative 
treatment effects ($d_k$); `eps`, which is a ridge penalty multiplier - this 
additional penalty is added when there is not enough data to estimate all 
parameters - the sparser the data, the more likely a larger ridge penalty is 
needed; `center`, an option to center the covariates; `rescale`, and option
to specify the desired scaling of covariates - this is especially important when
there are multiple covariates as they should all be converted to the same scale 
to ensure fair penalization.

In this example, we solve the GFL-NMR problem by penalizing the $d_k$'s and 
$\beta_{k,p}$'s equally by specifying `gamma = 1`. We also specify the ridge 
penalty multiplier to be `eps = 0.5`. We have kept the default `fit_full = FALSE` 
as there is not enough evidence to estimate all parameters in a full NMR model 
($\lambda = 0$). Also, since we are penalizing  one covariate, 
rescaling is not necessary.

```{r}
# Solve GFL-NMR problem
mod <- solve_gflnma(y = dat$TE, 
                    var_cov = var_cov, 
                    x_cov = x_cov, 
                    treatment1 = dat$treat1, 
                    treatment2 = dat$treat2, 
                    ref = "1",
                    fit_full = FALSE,
                    class1 = NULL, class2 = NULL, ref_class = NULL,
                    gamma = 1, 
                    eps = 0.5,
                    center = FALSE, 
                    rescale = FALSE)
```
In addition to outputting information on the solution path, `solve_gflnma()` 
outputs include `par_groups` which lists the parameter names and indices in the
solution path; the parameters names and indices are helpful inputs for other 
functions.

```{r}
mod$par_groups
```

Note "b1" refers to $\beta_{k,1}$, the interaction terms for the covariate in 
the first column of `x_cov`.

#### Visualize the solution path

We can plot the solution path using the `plot_solution_path()` function. At a 
minimum, this function requires the `solution` object outputted by 
`solve_gflnma()`. We additionally include the `par_groups` object outputted by 
`solve_gflnma()` to properly label our parameters.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   par_groups = mod$par_groups,
                   xlimits = c(-0.25, 6),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

Two plots are outputted for the two sets of parameters that are grouped 
separately: the $d_k$'s and the $\beta_{k,p}$'s.

The limits of x- and y-axis may be set to better visualize the path. This
may need to be done multiple times to get the desired scale for each plot.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   par_groups = mod$par_groups,
                   lambda_scale = 1,
                   xlimits = c(-0.05, 0.5),
                   #ylimits = c(-0.5, 0.5),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

Note that, these solution paths should be carefully inspected for extreme 
values, which may suggest a larger ridge penalty is needed.

To better visualize critical $\lambda$ values for which the $d_k$'s or 
$\beta_{k,p}$ pool or unpool, we need to run `tabulate_solution_path`, which is 
done in the next step.

### Select best fitting model
We can summarize the solution path with `tabulate_solution_path()`. 

```{r, attr.output='style="max-height: 100px;"'}
# Summarize GFL-NMR model fit statistics
(mod_sum <- tabulate_solution_path(mod$solution, 
                                   par_groups = mod$par_groups))
```

To select a sensible parameter pooling configuration (as determined by lambda 
$(\lambda)$), we recommend examining the AICc. The model that has the smallest 
AICc can be spotted by assessing whether `deltaAICc = 0`.

```{r}
best_mod_index <- which(mod_sum$mod_sum_df[, "deltaAICc"] == 0)
(best_groups <- mod_sum$mod_sum_df[best_mod_index, ]$PooledGroups)
```

In the model with the smallest AICc, $d_2,d_3,d_9$, $d_4,d_5, d_7,d_8,d_{10},d_{11},d_{12},d_{13},d_{14},d_{15},d_{16},d_{17}$ are pooled into 
two treatment groups and all $\beta_{k,1}$ have been shrunk to 0 (since they 
have been pooled with $\beta_{1,1}$), suggesting that regression is not needed. 
However, this is likely an artifact of penalizing the $d_k$'s the same way as 
the $\beta_{k,1}$'s (`gamma = 1`). As such, we recommend considering a range of 
`gamma`, as demonstrated later.

### Solve GFL-NMA problem: gamma = Inf

Note that it possible to penalize the interaction terms only by specifying 
`gamma = Inf` in `solve_gflnma()`. `Inf` implies we want to penalize the 
interaction terms infinitely more times than the relative effects.

```{r}
# Solve GFL-NMR problem
mod <- solve_gflnma(y = dat$TE, 
                    var_cov = var_cov, 
                    x_cov = x_cov, 
                    treatment1 = dat$treat1, 
                    treatment2 = dat$treat2, 
                    ref = "1",
                    fit_full = FALSE,
                    class1 = NULL, class2 = NULL, ref_class = NULL,
                    gamma = Inf, 
                    eps = 0.5,
                    center = FALSE, 
                    rescale = FALSE)
```

#### Visualize the solution path

Visually inspect the solution path for any extreme values which may suggest that
the ridge penalty multiplier needs to be increased.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   par_groups = mod$par_groups,
                   xlimits = c(-0.1, 1.5),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

Note that none of the $d_k$'s have shrunk to zero; their estimates only change
based on the shrunken values of $\beta_{k,1}$. 

### Select best fitting model
We can summarize the solution path with `tabulate_solution_path()`. 

```{r}
# Summarize GFL-NMR model fit statistics
(mod_sum <- tabulate_solution_path(mod$solution, 
                                   par_groups = mod$par_groups))
```

To select a sensible parameter pooling configuration (as determined by lambda 
$(\lambda)$), we recommend examining the AICc. The model that has the smallest 
AICc can be spotted by assessing whether `deltaAICc = 0`.

```{r}
best_mod_index <- which(mod_sum$mod_sum_df[, "deltaAICc"] == 0)
(best_groups <- mod_sum$mod_sum_df[best_mod_index, ]$PooledGroups)
```
In the model with the smallest AICc, all $\beta_{k,1}$ have been shrunk to 0 
(since they have been pooled with $\beta_{1,1}$), suggesting that regression is 
not needed. Nevertheless, we should still consider a decent range for `gamma`, 
which will be demonstrated shortly.

#### Visualize the solution path

Let's visualize critical $\lambda$ values along the solution path by making 
use of the `mod_sum_list` outputted from `tabulate_solution_path()`. Here, we 
have specified the AICc to be the penalized measure of fit for which we want to 
base our model selection on (e.g., `penfit = "AICc"`), which is represented with 
a red vertical dashed line.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   par_groups = mod$par_groups,
                   gfl_sum = mod_sum$mod_sum_list,
                   penfit = "AICc",
                   xlimits = c(-0.5, 2),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

## Multiple problems

We will now vary the ratio of penalty factors (`gamma`), as well as the ridge 
multiplier (`eps`), and use the penalized fit measures to determine the most
suitable model (and `gamma` and `eps` in parallel).

### Solve GFL-NMA problem: multiple gamma and eps

```{r}
gamma <- c(Inf, seq(0.5, 2, by=0.5))
eps <- c(0.01, 0.1, 0.5)

# Solve GFL-NMR problem
mod <- solve_gflnma(y = dat$TE, 
                    var_cov = var_cov, 
                    x_cov = x_cov, 
                    treatment1 = dat$treat1, 
                    treatment2 = dat$treat2, 
                    ref = "1",
                    fit_full = FALSE,
                    class1 = NULL, class2 = NULL, ref_class = NULL,
                    gamma = gamma, 
                    eps = eps,
                    center = FALSE, 
                    rescale = FALSE)
```

#### Visualize the solution path

As multiple GFL-NMR problems have been solved, there are multiple soluton paths
to consider. Let's hold off on visualizing the solution path for now, until
we determine the most suitable `gamma` and `eps`.

### Select best fitting model

Summarize the solution path with `tabulate_solution_path()`. 

```{r}
# Summarize GFL-NMR model fit statistics
(mod_sum <- tabulate_solution_path(mod$solution, 
                                   par_groups = mod$par_groups))
```

Notice how we have multiple outputs for each combination of `gamma` and 
`eps`.

To select a sensible parameter pooling configuration (as determined by lambda 
$(\lambda)$), let's examine the AICc. The model that has the smallest 
AICc can be spotted by assessing whether `deltaAICc = 0`. 

```{r}
best_mod_index <- which(mod_sum$mod_sum_df[, "deltaAICc"] == 0)
(best_groups <- mod_sum$mod_sum_df[best_mod_index, ]$PooledGroups)
```

In the model with the smallest AICc, $d_3,d_{17}$, $d_4,d_5,d_7,d_8,d_{10},d_{11},d_{12},d_{13},d_{14},d_{15},d_{16}$, and 
$d_9, d_{15}$ are pooled into three treatment groups. Most interaction terms
have been shrunk to 0 (i.e., those pooled with $\beta_{1,1}$), but 
$\beta_{3,1}, \beta_{11,1}$ have been pooled together as a non-zero interaction 
group and $\beta_{6,1}$ is a standalone non-zero interaction term.

#### Visualize the solution path

Let's now visualize critical $\lambda$ values along the solution path that 
contains the best model. We need to first find this solution path, and the
`find_best_gfl_soln()` helper function can help us with this.

```{r}
(best_model_index <- find_best_gfl_soln(mod_sum, "AICc"))
```

Now we can plot the solution path for the GFL-NMR problem through specifying
this problem index in the `mod_index` argument.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   mod_index = best_model_index, 
                   par_groups = mod$par_groups,
                   gfl_sum = mod_sum$mod_sum_list,
                   penfit = "AICc",
                   xlimits = c(-0.5, 5),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

### Recode parameters

To recode the treatments, we may use the `recode_gfl_params()` function. This 
requires 1) the data set for which we will use to re-run the NMR, 2) the 
estimated values for $d_k$ and $\beta_{k,p}$ for the selected model, and 3) the 
`par_groups` object outputted by `solve_gflnma()`. 

To obtain the estimated values for $d_k$ and $\beta_{k,p}$ for the selected 
model, we can extract these from the solution path outputted by 
`solve_gflnma()`. We first need to specify the $\lambda$ value for the selected 
model, rounded to 3 decimal places.

```{r}
best_lambda <- round(mod_sum$mod_sum_df[which(mod_sum$mod_sum_df[, "deltaAICc"] == 0), ]$lambda, 3)
```

Now we can extract the estimated values for $d_k$ and $\beta_{k,p}$ based on 
this $\lambda$.

```{r}
best_param_est <- mod$solution[[best_model_index]]$beta[, paste0(best_lambda)]
```
Now we can obtain the new treatment codes based on the selected GFL-NMR model.
Here, we add the codes to the data set outputted by `prep_gfl_data()`, as we 
will be using this data set to fit the NMR model.

```{r}
dat_recode <- recode_gfl_params(gfl_dat = dat, 
                                param_est = best_param_est, 
                                par_groups = mod$par_groups)

dat_recode
```

Note we will now estimate 6 relative treatment effects and 2 interaction terms 
(the third interaction term was shrunk to 0).

```{r}
# relative treatment effect groups
sort(unique(c(dat_recode$treat1_gfl, dat_recode$treat2_gfl)))
# interaction terms
sort(unique(c(dat_recode$treat1_beta1_gfl, dat_recode$treat2_beta1_gfl)))
```

### Re-run NMR with reconfigured parameters

We can now fit a multiplicative NMR model by coding the appropriate model in a 
Bayesian framework. It also possible to fit a multiplicative NMR model using the
`lm()` function - we will add a function to the `gflnma` package to enable this 
soon.

\
\

# Demonstration with multiple covariates

Coming soon!

\
\

# Note on fixed vs. random vs. multiplicative effects

The NMR model can be written in the form
$$ \mathbf{y} = \mathbf{XB} + \boldsymbol{\epsilon}  $$
where $\mathbf{y} = (\mathbf{y}_1, ..., \mathbf{y}_M)'$, $\mathbf{y}_m, m = 1, \dots, M$ 
are vectors of the $T_m$ relative treatment effects vs. a study-specific 
reference treatment estimated in each study $m$, $\mathbf{X}$ is
a design matrix that indicates the parameters in $\mathbf{B} = (\mathbf{d} ~ \boldsymbol{\beta}_1 \dots \boldsymbol{\beta}_p)$ 
being estimated in each study, and $\boldsymbol{\epsilon}$ is a multivariate 
normally distributed term with known covariance matrix. The covariance matrix is 
based on whether a fixed, random, or multiplicative effects model is assumed.

\

For a fixed effect model, 
$$ \mathbf{y}_m \sim \text{MVN}(\mathbf{XB}, \mathbf{V}_m) $$
where $\mathbf{V}_m$ is the variance-covariance matrix corresponding to the 
reported variances (and covariances) of the observed $\mathbf{y}_m$.

\

For a random effects model, 
$$ \mathbf{y}_m \sim \text{MVN}(\mathbf{XB}, \mathbf{\Omega}_m + \mathbf{V}_m) $$
where 
$$\mathbf{\Omega}_m = \begin{pmatrix}
         \tau^2 & 0.5\tau^2 & \cdots & 0.5\tau^2 \\
       0.5\tau^2 & \tau^2 & \cdots & 0.5\tau^2 \\
        \vdots & \vdots & \ddots & \vdots \\
        0.5\tau^2  & 0.5\tau^2 & \cdots & \tau^2
    \end{pmatrix}$$
and is of dimension $(T_m - 1) \times (T_m - 1)$.

\

For a multiplicative effects model, 
$$ \mathbf{y}_m \sim \text{MVN}(\mathbf{XB}, \phi \mathbf{V}_m) $$
such that that variance of the observed relative treatment effects are known
up to some constant $\phi$.

\
\

# References

Cooper NJ, Sutton AJ, Morris D, Ades AE, Welton NJ. Addressing between-study 
heterogeneity and inconsistency in mixed treatment comparisons: Application to 
stroke prevention treatments in individuals with non-rheumatic atrial 
fibrillation. Stat Med. 2009;28:1861-1881. doi:/10.1002/sim.3594

van Valkenhoef G. (2023). gemtc: Network Meta-Analysis Using Bayesian Methods. 
CRAN package. R package version 1.0-2. 
