---
title: "Demonstration of GFL-NMA via the `gflnma` package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GFL-NMA Example: Parkinson's}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gflnma)
library(netmeta)
```

# Data

To illustrate GFL-NMA, we will make use of a network of evidence on 5 treatments
for Parkinson's disease [Dias 2011]. There are 7 studies in this data set. The 
outcome reported in each treatment arm is the mean reduction in time off work. 
The relative effect measure (`TE`) is the mean difference.

Let's examine the data more closely. Note that, since the `gflnma` R package 
relies on the `netmeta` R package to compute the between-study standard 
deviation in random effects models, all $\binom{T_m}{2}$ pairwise contrasts in a 
multi-arm trial $m$ comparing $T_m$ treatments must be inputted (e.g., 3 
contrasts for 3-arm trials, 6 contrast for 4-arm trials, etc.). 

```{r}
# View data
dat_parkinsons
```
The only multi-arm trial in this data set is Guttman 1997. It compares three 
treatments, and thus we have three contrasts.

The data set contains arm-level (`mean1`, `sd1`, `n1`, `mean2`, `sd2`, `n2`) and 
contrast-level data (`TE`, `seTE`). While GFL-NMA requires contrast-level data,
we have included arm-level data which will later be used to compute the 
covariance between the relative effects in multi-arm trials.

## Inspect network

We can call upon `netconnection()` from the `netmeta` R package to assess if the 
network is connected and `netgraph()`, also from the `netmeta` R package, to 
plot the network diagram.

```{r, fig.dim = c(6, 6)}
# Check to see if network is connected
(netcon_check <- netmeta::netconnection(dat_parkinsons, 
                                        treat1 = treat1, 
                                        treat2 = treat2, 
                                        studlab = studlab))

# Plot the network diagram
netmeta::netgraph(netcon_check)
```

The output from `netmeta::netconnection()` tells us there is only one 
network, meaning there is a single connected network.

# Derive inputs for GFL-NMA

To implement the GFL-NMA framework, we require 1) a data frame that provides 
contrast-level data on relative treatment effects vs. a trial-specific reference 
treatment estimated in each study, and 2) the corresponding variance-covariance 
matrix. `prep_gfl_data()` will derive these outputs based on an inputted data 
set that contains all pairwise contrasts in multi-arm studies. 

The outputted variance-covariance matrix depends on whether a fixed, random, or
multiplicative effects model is desired, which must be specified by the 
`modtype` argument. See note at the end of this vignette for more details.

```{r}
dat_list <- prep_gfl_data(dat = dat_parkinsons, 
                          TE = "TE",
                          seTE = "seTE", 
                          treatment1 = "treat1", 
                          treatment2 = "treat2",
                          studlab = "studlab",
                          sd1 = "sd1", 
                          n1 = "n1",
                          modtype = "FE")

(dat <- dat_list$dat)

(var_cov <- dat_list$var_cov)

(mod_type <- dat_list$mod_type)
```

Notice how the data in the multi-arm study (Guttman 1997) only contains evidence
on the relative treatment effects vs. placebo.

# Solve GFL-NMA problem

Now that we have the required data inputs, we may solve the GFL-NMA problem via
the `solve_gflnma()` function. In the code below, we have set `fit_full = TRUE` 
to indicate that the full NMA model (penalty factor $\lambda = 0$) should be 
considered when we compare the fit of multiple models (with different treatment
configurations) along the solution path.

```{r}
# Solve GFL-NMA problem
mod <- solve_gflnma(y = dat$TE, 
                    var_cov = var_cov, 
                    treatment1 = dat$treat1, 
                    treatment2 = dat$treat2, 
                    ref = "Placebo",
                    fit_full = TRUE)
```

In addition to outputting information on the solution path, `solve_gflnma()` 
outputs include `par_groups` which lists the parameter names and indices in the
solution path; the parameters names and indices are helpful inputs for other 
functions.

```{r}
mod$par_groups
```

## Visualize the solution path

We can plot the solution path using the `plot_solution_path()` function. At a 
minimum, this function requires the `solution` object outputted by 
`solve_gflnma()`. We additionally include the `par_groups` object outputted by 
`solve_gflnma()` to properly label our parameters.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   par_groups = mod$par_groups,
                   xlimits = c(-0.25, 5),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

To better visualize critical $\lambda$ values for which the $d_k$'s pool or 
unpool, we need to run `tabulate_solution_path`, which is done in the next step.

# Select best fitting model

Along the solution path computed by `solve_gflnma()`, there are multiple points
where the relative treatment effect parameters $(d_k)$ pool and unpool. We want 
to compare the fit of these models to select the most appropriate model. 
`tabulate_solution_path()` computes several penalized model fit statistics for
each of these models: the Akaike information criterion (AIC), the AIC corrected
for small sample size (AICc), the Bayesian information criterion (BIC), the BIC
corrected for small sample size (BICc). deltaAIC, deltaAICc, deltaBIC, and 
deltaBICc present the computed respective difference between the AIC, AICc, BIC, 
and BICc of a given model and the model that has the smallest penalized fit.

To compute these model fit statistics, `tabulate_solution_path()` expects the 
`solution` and `par_groups` objects outputted by `solve_gflnma()`. If the full 
model can be estimated, `tabulate_solution_path()` expects the residual sum of 
squares (RSS) for the full model to be inputted as well, and this may be 
obtained from the `full_RSS` object outputted by `solve_gflnma()`.

```{r}
# Summarize GFL-NMA model fit statistics
(mod_sum <- tabulate_solution_path(mod$solution, 
                                   par_groups = mod$par_groups, 
                                   full_rss = mod$full_RSS))
```

To select a sensible parameter pooling configuration (as determined by lambda 
$(\lambda)$), we recommend examining the AICc. The model that has the smallest 
AICc can be spotted by assessing whether `deltaAICc = 0`.

```{r}
best_mod_index <- which(mod_sum$mod_sum_df[, "deltaAICc"] == 0)
(best_groups <- mod_sum$mod_sum_df[best_mod_index, ]$PooledGroups)
```

In the model with the smallest AICc, $d_2$ and $d_5$ are pooled.

## Visualize the solution path

We may now visualize critical $\lambda$ values along the solution path by making 
use of the `mod_sum_list` outputted from `tabulate_solution_path()`. We can also 
specify the penalized measure of fit for which we want to base our model 
selection on (e.g., `penfit = "AICc"`), which will then be represented with a 
red vertical dashed line.

```{r, fig.dim = c(7, 6)}
plot_solution_path(gfl_soln = mod$solution, 
                   par_groups = mod$par_groups,
                   gfl_sum = mod_sum$mod_sum_list,
                   penfit = "AICc",
                   xlimits = c(-0.25, 5),
                   cex_lab = 2, cex_axis = 1.5, cex_text = 1)
```

# Recode treatments

To recode the treatments, we may use the `recode_gfl_params()` function. This 
requires 1) the data set for which we will use to re-run the NMA, 2) the 
estimated values for $d_k$ for the selected model, and 3) the `par_groups`
object outputted by `solve_gflnma()`. Additionally, if the GFL-NMA problem was
solved based on a data set that only included the character names, we need to 
input 4) a data frame consisting of the treatment names and codes.

To obtain the estimated values for $d_k$ for the selected model, we can extract
these from the solution path outputted by `solve_gflnma()`. We first need to 
specify the $\lambda$ value for the selected model, rounded to 3 decimal places.

```{r}
best_lambda <- round(mod_sum$mod_sum_df[which(mod_sum$mod_sum_df[, "deltaAICc"] == 0), ]$lambda, 3)
```

Now we can extract the estimated values for $d_k$ based on this $\lambda$.

```{r}
best_param_est <- mod$solution[[1]]$beta[, paste0(best_lambda)]
```

Since the GFL-NMA problem was run on `dat` which only contains the treatment 
names as character strings, we must construct a data frame consisting of the 
treatment names and codes.

```{r}
trt_codes <- data.frame(code = 1:length(unique(c(dat$treat1, dat$treat2))),
                        trt = c("Placebo", "Bromocriptine", "Cabergoline", 
                                "Pramipexole", "Ropinirole"))
```

Now we can obtain the new treatment codes based on the selected GFL-NMA model.

```{r}
dat_recode <- recode_gfl_params(gfl_dat = dat_parkinsons, 
                                param_est = best_param_est, 
                                par_groups = mod$par_groups,
                                trt_codes = trt_codes)

dat_recode
```

# Re-run NMA with recongifured treatments

Now, we may re-run the NMA based on the new treatment codes. Here, we will run a 
frequentist NMA via the `netmeta()` function from the `netmeta` package. 

To label the synthesized relative effects outputted by `netmeta::netmeta()`, we
have opted to label the new treatment codes first.

```{r}
# Label recoded treatments
gfl_codes <- data.frame(code = 1:length(unique(c(dat_recode$treat1_gfl, dat_recode$treat2_gfl))),
                        trt = c("Placebo", "Bromocriptine or Ropinirole", "Cabergoline", "Pramipexole"))
dat_recode$treat1_gfl <- factor(dat_recode$treat1_gfl, levels = gfl_codes$code, labels = gfl_codes$trt)
dat_recode$treat2_gfl <- factor(dat_recode$treat2_gfl, levels = gfl_codes$code, labels = gfl_codes$trt)
```

Note that `netmeta::netmeta()` does not accept constrasts between the same 
treatments, so contrasts between pooled treatments need to be removed.

```{r}
# Drop rows that compare the same pooled treatments
dat_recode <- dat_recode[-which(dat_recode$treat1_gfl == dat_recode$treat2_gfl), ]
```

```{r}
# Run NMA
net_postgfl <- netmeta::netmeta(TE = TE, 
                                seTE = seTE, 
                                treat1 = treat1_gfl, 
                                treat2 = treat2_gfl, 
                                studlab = studlab,
                                data = dat_recode, 
                                sm = "MD", 
                                reference = "Placebo",
                                common = TRUE,
                                random = FALSE)

net_postgfl
```

\
\

# Note on fixed vs. random vs. multiplicative effects

The NMA model can be written in the form
$$ \mathbf{y} = \mathbf{Xd} + \boldsymbol{\epsilon}  $$
where $\mathbf{y} = (\mathbf{y}_1, ..., \mathbf{y}_M)'$, $\mathbf{y}_m, m = 1, \dots, M$ 
are vectors of the $T_m$ relative treatment effects vs. a study-specific 
reference treatment estimated in each study $m$, $\mathbf{X}$ is
a design matrix that indicates the relative treatment effects $\mathbf{d}$ 
being estimated in each study, and $\boldsymbol{\epsilon}$
is a multivariate normally distributed term with known covariance matrix. The
covariance matrix is based on whether a fixed, random, or multiplicative effects
model is assumed.

\

For a fixed effect model, 
$$ \mathbf{y}_m \sim \text{MVN}(\mathbf{Xd}, \mathbf{V}_m) $$
where $\mathbf{V}_m$ is the variance-covariance matrix corresponding to the 
reported variances (and covariances) of the observed $\mathbf{y}_m$.

\

For a random effects model, 
$$ \mathbf{y}_m \sim \text{MVN}(\mathbf{Xd}, \mathbf{\Omega}_m + \mathbf{V}_m) $$
where 
$$\mathbf{\Omega}_m = \begin{pmatrix}
         \tau^2 & 0.5\tau^2 & \cdots & 0.5\tau^2 \\
       0.5\tau^2 & \tau^2 & \cdots & 0.5\tau^2 \\
        \vdots & \vdots & \ddots & \vdots \\
        0.5\tau^2  & 0.5\tau^2 & \cdots & \tau^2
    \end{pmatrix}$$
and is of dimension $(T_m - 1) \times (T_m - 1)$.

\

For a multiplicative effects model, 
$$ \mathbf{y}_m \sim \text{MVN}(\mathbf{Xd}, \phi \mathbf{V}_m) $$
such that that variance of the observed relative treatment effects are known
up to some constant $\phi$.

\
\

# References

Dias S, Welton NJ, Sutton AJ, Ades AE. NICE DSU Technical Support Document 2: A 
generalised linear modelling framework for pairwise and network meta-analysis of 
randomised controlled trials. 2011; last updated September 2016; available from 
http://www.nicedsu.org.uk

Kong X, Daly CH, Béliveau A. Generalized fused lasso for treatment pooling in 
network meta-analysis. Stat Med. 2024;43(30):5635-5649. doi:10.1002/sim.10253


